### 2015-07-15 Paper_v2.tex
New Updates

The results of the model suggest that an inhibition of either the PI3K pathway or the MAPK pathway in combination with an AR inhibitor as a possible therapy for CRPC. 
Sensitivity analysis revealed no change in the top sensitive species in the presence or absence of the AR inhibitor, enzalutamide. 
PI3K/AKT and MAPK species continued to fall in the top two percent of sensitive species. 
A study by Carver \textit{et al.} looked at dual inhibition of AR and PI3K signaling in LNCaP cells and in a Pten-deficient murine prostate cancer model \cite{Parker2013}. 
Using both the PI3K inhibitor, BEZ235, and the AR inhibitor, MDV3100 (enzalutamide), the group saw a drastic decrease in the total number of cells. 
Each inhibitor on it’s own had a much smaller effect on total cell number. They saw an increase in the cell death marker, c-PARP, in the dual inhibition case. 
The group hypothesized that AKT inhibition leads to increased AR signaling activity through increased protein concentrations of HER3. 
On the other hand AR inhibition leads to increased AKT activity due to the down regulation of PHLPP, a protein phosphatase that regulates AKT. 
For the simplicity of this model, the HER3 pathway and also cell death were not included in the model. 
Dual knock-out studies of PI3K and AR in our model show no additive effect on any cell cycle proteins through the dual knock-out. 
The pathways in our model appear to be uncoupled and therefore no synergy is shown in the dual knock-out case. 
This could indicate that the combined decrease in cell population is entirely due to cell death. 
The Carver \textit{et al.} study did not consider cell cycle proteins or cell growth. 
Our model does show a decrease in cell cycle proteins in the PI3K knock-out as well as in the dual knock-out case. 
This result seems to be consistent with the decreased cell growth in the PI3K knock-out case which is not dependent on cell death, as c-PARP levels are low. 
The decrease in cell cycle proteins is due to a decrease in general translation, including free eIF4E levels and activated 40S ribosome subunit. 
The decrease in p70 (S6) activation due to inhibition of PI3K is shown in both the model and in the Carver \textit{et al.} study ((Fig. \ref{fg:Supp_Figure_KO_new.pdf})), indicating this result is due to the PI3K pathway. 
Our results confirm the Carver \textit{et al.} study in that a dual inhibition of AR and PI3K signaling led to a more prominent decrease in cell viability then each of the inhibitors alone. 
We extended the study to look at the addition of a third inhibitor, sorafenib, that inhibits Raf kinase in the MAPK pathway.

@@

\subsection*{Cell culture and treatments}
Androgen dependent LNCaP prostate cancer cells were a gift from Dr. Brian Kirby (Cornell University), and the Castrate Resistant C4-2 prostate cancer cell line was purchased from MD Anderson Cancer Center, University of Texas. Cell lines were maintained in RPMI 1640 media (Life Technologies, Inc., Grand Island, NY) with 10\% fetal calf serum (FBS; Hyclone) and 1x antibiotic/antimycotic (Sigma, St. Louis, MO) in a 5\% CO$_2$ humidified atmosphere at 37$^{\circ}$C. The AR inhibitor MDV3100 (enzalutamide) and the Raf inhibitor sorafenib were purchased from SantaCruz Biotechnology (Santa Cruz, CA). The PI3K inhibitor LY294002 was purchased from Cell Signaling Technologies (Danvers, MA, USA). All stock solutions were diluted in DMSO (Sigma, St. Louis, MO).

@@

@@

\begin{figure}\centering
\includegraphics[width=1.0\textwidth]{./figs/Fig_6_PI3K_AR_KO_3_Cases.pdf}
\caption{Robustness analysis of a population of CR prostate models with PI3K (blue), AR (green), and PI3K and AR (orange) knock-outs (N = 500). A log fold change of greater then zero implies that the concentration of the protein increased with the knock-out, while a log fold change of less then zero indicates that the concentration of protein decreased. A log of fold change equal to 0, shows no response due to the knock-out.
 A.,B.,C. Log robustness of cyclin E, PSA, and p70p versus Cyclin D for the three knock-out cases. A CR LNCaP cell was assumed for all knock-out cases.}
\label{fg:Supp_Dual_Knockout}
\end{figure}

>>>>

The results of the model suggest that an inhibition of either the PI3K pathway or the MAPK pathway in combination with an AR inhibitor as a possible therapy for CRPC. 
Sensitivity analysis revealed no change in the top sensitive species in the presence or absence of the AR inhibitor, enzalutamide. 
PI3K/AKT and MAPK species continued to fall in the top two percent of sensitive species. 
A study by Carver \textit{et al.} looked at dual inhibition of AR and PI3K signaling in LNCaP cells and in a Pten-deficient murine prostate cancer model \cite{Parker2013}. 
Using both the PI3K inhibitor, BEZ235, and the AR inhibitor, MDV3100 (enzalutamide), the group saw a drastic decrease in the total number of cells. 
Each inhibitor on it’s own had a much smaller effect on total cell number. They saw an increase in the cell death marker, c-PARP, in the dual inhibition case. 
The group hypothesized that AKT inhibition leads to increased AR signaling activity through increased protein concentrations of HER3. 
On the other hand AR inhibition leads to increased AKT activity due to the down regulation of PHLPP, a protein phosphatase that regulates AKT. 
For the simplicity of this model, the HER3 pathway and also cell death were not included in the model. 
Dual knock-out studies of PI3K and AR in our model show only a slight additive effect on cell cycle proteins through the dual knock-out (Fig. \ref{fg:Dual_Knockout}).
The pathways in our model appear to be uncoupled and therefore little synergy is shown in the dual knock-out case. 
This could indicate that the combined decrease in cell population due to the dual inhibition of PI3K and AR is entirely due to cell death. 
The Carver \textit{et al.} study did not consider cell cycle proteins or cell growth. 
Our model does show a decrease in cell cycle proteins in the PI3K knock-out as well as in the dual knock-out case. 
This result seems to be consistent with the decreased cell growth in the PI3K knock-out case which is not dependent on cell death, as c-PARP levels are low. 
The decrease in cell cycle proteins in the model is due to a decrease in general translation, including free eIF4E levels and activated 40S ribosome subunit. 
The decrease in p70 (S6) activation due to inhibition of PI3K is shown in both the model and in the Carver \textit{et al.} study , indicating this result is due to the PI3K pathway (Fig. \ref{fg:Supp_KO}). 
Our experimental results confirm the Carver \textit{et al.} study in that a dual inhibition of AR and PI3K signaling led to a more prominent decrease in cell viability then each of the inhibitors alone in LNCaP cells. 
We extended the study to look at the addition of a third inhibitor, sorafenib, that inhibits Raf kinase in the MAPK pathway and an additional cell line, C4-2, which was CR.  In both cell lines, there was not a significant decrease in cell viability between the three dual inhibitor cases and the triple inhibition case at 74 hours.

@@

\subsection*{Cell culture and treatments}
Androgen dependent LNCaP prostate cancer cells were a gift from Dr. Brian Kirby (Cornell University), and the castration resistant C4-2 prostate cancer cell line was purchased from MD Anderson Cancer Center, University of Texas. Cell lines were maintained in RPMI 1640 media (Life Technologies, Inc., Grand Island, NY) with 10\% fetal calf serum (FBS; Hyclone) and 1x antibiotic/antimycotic (Sigma, St. Louis, MO) in a 5\% CO$_2$ humidified atmosphere at 37$^{\circ}$C. The AR inhibitor MDV3100 (enzalutamide) and the Raf inhibitor sorafenib were purchased from SantaCruz Biotechnology (Santa Cruz, CA). The PI3K inhibitor LY294002 was purchased from Cell Signaling Technologies (Danvers, MA, USA). All stock solutions were diluted in DMSO (Sigma, St. Louis, MO).

@@

\begin{figure}\centering
\includegraphics[width=1.0\textwidth]{./figs/Fig_8_PI3K_AR_KO_3_Cases.pdf}
\caption{Robustness analysis of a population of CR prostate models with PI3K (blue), AR (green), and PI3K and AR (orange) knock-outs (N = 500). A log fold change of greater then zero implies that the concentration of the protein increased with the knock-out, while a log fold change of less then zero indicates that the concentration of protein decreased. A log of fold change equal to 0, shows no response due to the knock-out.
 A.,B.,C. Log robustness of cyclin E, PSA, and p70p versus Cyclin D for the three knock-out cases. A CR LNCaP cell was assumed for all knock-out cases.}
\label{fg:Dual_Knockout}
\end{figure}

@@
